Real Time Variational Density Forecasting

### 2017-03-01 Confirmation/confirmation_V2.tex
Started work on slides

The literature on Variational Bayes is dominated by two major implementations: Mean Field Variational Bayes (MFVB) and what we call Stochastic Variational Bayes (SVB). For either implementation we emphasise that there are two seperate tasks: Choose a functional form for the approximating distribution, and then optimise the parameters of this distribution to minimise a well defined divergence between the approximation and the true posterior.

@@

We illustrate the implementation of MFVB with the model considered in Section (\ref{sec:MCMC}), where $y_t$ for $t = 1, \dots, T$ is generated independently from a $\mathcal{N}(\mu, \sigma^2)$ distribution, and prior distributions are given by $\mu \sim \mathcal{N}(\gamma, \sigma^2 / \tau)$ and $\sigma^2 \sim IG(\mbox{shape} = \alpha, \mbox{scale =}\beta)$. It can be shown that
\begin{equation}
\label{mf3}
q(\mu | \boldsymbol{\lambda}_1) \propto \exp \left\{ \frac{-(T + \tau) \mathbb{E}_{q(\sigma^2)}(\sigma^{-2})}{2} \left( \mu - \frac{\tau \gamma + \sum_{t=1}^{T} y_t}{T + \tau} \right)^2 \right\}
\end{equation}
and
\begin{align}
\label{mf4}
q(\sigma^2 | \boldsymbol{\lambda}_2) &\propto \sigma^{-2((T+1)/2 + \alpha + 1)} \nonumber \\
\times &\exp \left\{ \frac{ \beta + 1/2\left((\tau + T)\mathbb{E}_{q(\mu)}(\mu^2) - 2 \mathbb{E}_{q(\mu)}(\mu)(\sum_{t=1}^{T}y_t + \tau) + \sum_{t=1}^{T} y_t^2 + \tau \gamma^2\right)}{-\sigma^2} \right\}.
\end{align}

As the prior distributions used were conjugate to the likelihood, the Variational Bayes optimal distributions for $q(\mu |  \boldsymbol{\lambda}_1)$ and $q(\sigma^2 | \boldsymbol{\lambda}_2)$ are the same family as the prior distributions, and are denoted by $q(\mu | \boldsymbol{\lambda}_1) \sim \mathcal{N}(\tilde{\gamma}, \tilde{\tau})$ and $q(\sigma^2 | \boldsymbol{\lambda}_2) \sim IG(\tilde{\alpha}, \tilde{\beta})$. The vector $\boldsymbol{\lambda} = (\tilde{\gamma}, \tilde{\tau}, \tilde{\alpha}, \tilde{\beta})'$ is given by
\begin{align}
\tilde{\gamma} &= \frac{\tau \gamma + \sum_{t=1}^{T} y_t}{T + \tau}  \label{mf5} \\ 
\tilde{\tau} &= \left((T + \tau)\mathbb{E}_{q(\sigma^2)}(\sigma^{-2})\right )^{-1} \label{mf6} \\
\tilde{\alpha} &= (T+1)/2 + \alpha  \label{mf7} \\
\tilde{\beta} &= \beta + 1/2\left((\tau + T)\mathbb{E}_{q(\mu)} - 2 \mathbb{E}_{q(\mu)}(\mu)(\sum_{t=1}^{T}y_t + \tau) + \sum_{t=1}^{T} y_t^2 + \tau \gamma^2 \right). \label{mf8}
\end{align}.

The expectations in (\ref{mf5}) - (\ref{mf8}) are available in a closed form for these distributions, and these are substituted in to derive the set of mean field equations:
\begin{align}
\tilde{\tau} &= \frac{\tilde{\beta}}{\tilde{\alpha}(T + \tau)} \label{mf9} \\
\tilde{\beta} &= \beta + 1/2\left((\tau + T)(\tilde{\gamma}^2 + \tilde{\tau}) - 2 \tilde{\gamma}(\sum_{t=1}^{T}y_t + \tau) + \sum_{t=1}^{T} y_t^2 + \tau \gamma^2 \right)\label{mf10}
\end{align}
$\tilde{\gamma}$ and $\tilde{\alpha}$ are omitted from the mean field equations as their values do not depend on other $\boldsymbol{\lambda}$ parameters.

@@

almost everywhere, where $\epsilon \sim U(0, 1)$ and $p^{-1}$ denotes the inverse CDF of the true posterior, will yeild $q(\boldsymbol{\theta} | \boldsymbol{\lambda} = p(\boldsymbol{\theta} | y_{1:T}$ and thus minimise the KL divergence. Attempts to model a suitably flexible function $f(\cdot, \cdot)$ include heirarchical models \citep{Ranganath2016a}, neural networks (\citealp{Kingma2014}; \citealp{Rezende2014}; \citealp{Rezende2015}), and Gaussian processes \citep{Tran2016}.

@@

\end{document}
\grid
\grid
\grid
\grid
\grid

>>>>

The literature on Variational Bayes is dominated by two major implementations: Mean Field Variational Bayes (MFVB) and what we call Stochastic Variational Bayes (SVB). For either implementation we emphasise that there are two seperate tasks: Choose a functional form for the approximating distribution, and then optimise the parameters of this distribution to minimise a well defined divergence between the approximation and the true posterior. A recent review of Variational Bayes is provided by \citet{Blei2017}.

@@

We illustrate the implementation of MFVB with the model considered in Section (\ref{sec:MCMC}), where $y_t$ for $t = 1, \dots, T$ is generated independently from a $\mathcal{N}(\mu, \sigma^2)$ distribution, and prior distributions are given by $\mu \sim \mathcal{N}(\gamma, \sigma^2 / \tau)$ and $\sigma^2 \sim IG(\mbox{shape} = \alpha, \mbox{scale =}\beta)$. It can be shown that
\begin{equation}
\label{mf3}
q(\mu | \boldsymbol{\lambda}_1) \propto \exp \left\{ \frac{-(T + \tau) \mathbb{E}_{q(\sigma^2)}[\sigma^{-2}]}{2} \left( \mu - \frac{\tau \gamma + \sum_{t=1}^{T} y_t}{T + \tau} \right)^2 \right\}
\end{equation}
and
\begin{align}
\label{mf4}
q(\sigma^2 | \boldsymbol{\lambda}_2) &\propto \sigma^{-2((T+1)/2 + \alpha + 1)} \nonumber \\
\times &\exp \left\{ \frac{ \beta + 1/2\left((\tau + T)\mathbb{E}_{q(\mu)}[\mu^2] - 2 \mathbb{E}_{q(\mu)}[\mu ]\left(\sum_{t=1}^{T}y_t + \tau \right) + \sum_{t=1}^{T} y_t^2 + \tau \gamma^2\right)}{-\sigma^2} \right\}.
\end{align}

As the prior distributions used were conjugate to the likelihood, the Variational Bayes optimal distributions for $q(\mu |  \boldsymbol{\lambda}_1)$ and $q(\sigma^2 | \boldsymbol{\lambda}_2)$ are the same family as the prior distributions, and are denoted by $q(\mu | \boldsymbol{\lambda}_1) \sim \mathcal{N}(\tilde{\gamma}, \tilde{\tau})$ and $q(\sigma^2 | \boldsymbol{\lambda}_2) \sim IG(\tilde{\alpha}, \tilde{\beta})$. The vector $\boldsymbol{\lambda} = (\tilde{\gamma}, \tilde{\tau}, \tilde{\alpha}, \tilde{\beta})'$ is given by
\begin{align}
\tilde{\gamma} &= \frac{\tau \gamma + \sum_{t=1}^{T} y_t}{T + \tau}  \label{mf5} \\ 
\tilde{\tau} &= \left((T + \tau)\mathbb{E}_{q(\sigma^2)}[\sigma^{-2}]\right )^{-1} \label{mf6} \\
\tilde{\alpha} &= (T+1)/2 + \alpha  \label{mf7} \\
\tilde{\beta} &= \beta + 1/2\left((\tau + T)\mathbb{E}_{q(\mu)}[\mu^2 ]- 2 \mathbb{E}_{q(\mu)}[\mu ]\left(\sum_{t=1}^{T}y_t + \tau\right) + \sum_{t=1}^{T} y_t^2 + \tau \gamma^2 \right). \label{mf8}
\end{align}.

The expectations in (\ref{mf5}) - (\ref{mf8}) are available in a closed form for these distributions, and these are substituted in to derive the set of mean field equations:
\begin{align}
\tilde{\tau} &= \frac{\tilde{\beta}}{\tilde{\alpha}(T + \tau)} \label{mf9} \\
\tilde{\beta} &= \beta + 1/2\left((\tau + T)(\tilde{\gamma}^2 + \tilde{\tau}) - 2 \tilde{\gamma}\left(\sum_{t=1}^{T}y_t + \tau\right) + \sum_{t=1}^{T} y_t^2 + \tau \gamma^2 \right)\label{mf10}
\end{align}
$\tilde{\gamma}$ and $\tilde{\alpha}$ are omitted from the mean field equations as their values do not depend on other $\boldsymbol{\lambda}$ parameters.

@@

almost everywhere, where $\epsilon \sim U(0, 1)$ and $p^{-1}$ denotes the inverse CDF of the true posterior, will yeild $q(\boldsymbol{\theta} | \boldsymbol{\lambda} = p(\boldsymbol{\theta} | y_{1:T}$ and thus minimise the KL divergence. Attempts to model a suitably flexible function $f(\cdot, \cdot)$ include heirarchical models \citep{Ranganath2016a}, neural networks (\citealp{Kingma2014}; \citealp{Rezende2014}; \citealp{Rezende2015}; \citealp{Kingma2016}), and Gaussian processes \citep{Tran2016}.

@@

\end{document}
\grid
\grid
\grid
\grid
\grid
\grid
\grid
